---
title: "Data Visualisation & Story Telling - Group Project - Spotify"
author: "Study Group 14"
date: 03/12/2020
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
devtools::install_github("ricardo-bion/ggradar", dependencies=TRUE)
```

```{r loading packages, include=FALSE}
#devtools::install_github("ricardo-bion/ggradar", dependencies = TRUE)
library(Rspotify)
library(ggplot2)
library(tidyverse)
library(janitor)
library(dplyr)
library(skimr)
library(fmsb) # for radar charts
library(ggradar)
library(ggridges)
library(ggiraph)
library(ggiraphExtra)
library(reshape2)
library(moonBook)
library(sjmisc)
library(gridExtra)
library(magick) # for gifs
library(tm)           # TEXT MINING LIBRARY R
library(SnowballC)    # R interface to the C 'libstemmer' stemming library
library(RColorBrewer) # colour pallets for wordcloud
library(wordcloud)    # standard wordcloud
library(wordcloud2)   # fancy wordcloud 
library(rpart.plot)
library(caret)
library(lubridate)
library(Hmisc)
library(readr)
library(gghalves)
#Spotify colours: #1db954(green) and #191414(black) for the Tableau
```

# Introduction
Have you listened to music today? Probably! Music is so present in everyone's life and we're constantly exposed to it everyday, all day.

**But have you ever thought about the data behind the music that you listen to every day?**

Chances are you you have used Spotify before. Did you know you could retrieve specific song, album and artist metadata from their Spotify Data Catalouge? Let's have a look at the data behind music on Spotify, one of the biggest music streaming services worldwide. We are going to analyze the audio features, content, time, location, and much more.

**Let us show you music in a completely different light and find out what makes popular music!**

## Spotify API set-up
The tool that we use to retrieve the data is the Spotify API for developers. It allows us to retrieve JSON metadata about music artists, albums, and tracks, directly from the Spotify Data Catalogue.

The instructions we referred to for setting up Spotify API are [here](https://r-music.rbind.io/posts/2018-10-01-rspotify/). You can find more information on the API or set it up yourself on [Spotify for Developers](https://developer.spotify.com). 

```{r set up spotify API}
# set up API connection with personal app_id, client_id, client_secret
# feel free to connect your own account!

# keys <- spotifyOAuth("app_id","client_id","client_secret")
keys <- spotifyOAuth("dataviz","b9bd44a451e94288af529470c5d3ccf0", "c0f6d90ce49d4ded95887bf45c46f958")

# We wil use this key whenever we try to retrieve music from the API
```

```{r check API}
# Using  my own username
user <- getUser("sedroc2015", token = keys)
user$display_name # user name

# check followers
user$followers # :'( no followers
```

## Data

The audio features we are going to use for analysis are in the following three aspects.

*Mood:*
- Danceability: whether is song is danceable, calculated based on combination of tempo, beat, etc
- Valence: tracks with high valence sound more positive (happy and cheerful)
- Energy: whether the music is intense and active 
- Tempo: whether the overall beats per minute (BPM) is fast or slow

*Properties:*
- Loudness: loud music is usually high in decibels (dB)
- Speechiness: whether there is presence of spoken works (audio books have a high speechiness)
- Instrumentalness: whether the music has a lot of instruments or just vocals

*Context*
- Liveness: detects the presence of audience (more audience higher liveness)
- Acousticness: whether the music makes natural sound, instead of being produced by electrical equipment

# Our own music preferences

Music is heard by people daily in many parts of the world, and affects people in various ways from emotion regulation to cognitive development, along with providing a means for self-expression.

What we listen tells a lot about who we are. Here we analyze our Spotify play lists to find out the similarities and differences among us five.

# Radar charts

## Prepare data

### Marie

Marie's Top Songs 2020

- original: https://open.spotify.com/playlist/37i9dQZF1EM22Em6vPrOSO

Retrieve playlist through API, then retrieve the features of every song with a for loop and create a data set with all the songs and their features

https://open.spotify.com/playlist/37i9dQZF1EM22Em6vPrOSO

```{r retrieve playlist}
# get playlist
marie_top2020_basic <- getPlaylistSongs("spotify","37i9dQZF1EM22Em6vPrOSO", token =  keys)

# get all features for that playlist

# create data set with any 1 song
marie_top2020_wfeat <- getFeatures("3bH4HzoZZFq8UpZmI2AMgV", token = keys)

# for loop all song ids from the playlist through the getFeatures function
for (i in marie_top2020_basic$id){
  marie_top2020_wfeat <- rbind(marie_top2020_wfeat, getFeatures(i, token = keys))
}
                  
```

Fix playlist data set and delete duplicates

```{r fix data}
# get rid of first line (that we added to set up the data frame)
marie_top2020_wfeat <- marie_top2020_wfeat[-1,]

# join playlist df and feature df to have all song info in one data set
marie_top2020 <- left_join(marie_top2020_basic, marie_top2020_wfeat, by = "id") 

# fix data set
marie_top2020 <- marie_top2020 %>% 
                  # eliminate irrelevant columns
                  select(-uri, -analysis_url, -time_signature) %>% 
                  # add 1 columns with name for later when joining all of us 
                  cbind("Marie") %>% 
                  # rename that column
                  rename(name = '"Marie"')
```

```{r clean data}
# check if there are duplicates
dupes <- marie_top2020 %>%
  get_dupes(id)

# get rid of duplicates
marie_top2020 <- marie_top2020 %>%
  unique()

# check
dupes <- marie_top2020 %>%
  get_dupes(id)

# check data 
#head(marie_top2020)
#glimpse(marie_top2020)
#skim(marie_top2020)

```

> Now repeat for everyone else

### Sammy

- original: https://open.spotify.com/playlist/3mkjOvN3TUeazOwk59Jl7X

```{r retrieve playlist Sammy}
# get playlist
sammy_top2020_basic <- getPlaylistSongs("spotify","3mkjOvN3TUeazOwk59Jl7X", token =  keys)

# get all features for that playlist

# create data set with any 1 song
sammy_top2020_wfeat <- getFeatures("3bH4HzoZZFq8UpZmI2AMgV", token = keys)

# for loop all song ids from the playlist through the getFeatures function
for (i in sammy_top2020_basic$id){
  sammy_top2020_wfeat <- rbind(sammy_top2020_wfeat, getFeatures(i, token = keys))
}
                  
# get rid of first line (that we added to set up the data frame)
sammy_top2020_wfeat <- sammy_top2020_wfeat[-1,]

# join playlist df and feature df to have all song info in one data set
sammy_top2020 <- left_join(sammy_top2020_basic, sammy_top2020_wfeat, by = "id") 

# fix data set
sammy_top2020 <- sammy_top2020 %>% 
                  # eliminate irrelevant columns
                  select(-uri, -analysis_url, -time_signature) %>% 
                  # add 1 columns with name for later when joining all of us 
                  cbind("Sammy") %>% 
                  # rename that column
                  rename(name = '"Sammy"')

# check if there are duplicates
dupes <- sammy_top2020 %>%
  get_dupes(id)

# get rid of duplicates
sammy_top2020 <- sammy_top2020 %>%
  unique()

# check
dupes <- sammy_top2020 %>%
  get_dupes(id)

# check data 
#head(sammy_top2020)
#glimpse(sammy_top2020)
#skim(sammy_top2020)

```


### Ruchen

- original: https://open.spotify.com/playlist/3ncHsEr99JsKoZzgvE6EJ9

```{r retrieve playlist Ruchen}
# get playlist
ruchen_top2020_basic <- getPlaylistSongs("spotify","3ncHsEr99JsKoZzgvE6EJ9", token =  keys)

# get all features for that playlist

# create data set with any 1 song
ruchen_top2020_wfeat <- getFeatures("3bH4HzoZZFq8UpZmI2AMgV", token = keys)

# for loop all song ids from the playlist through the getFeatures function
for (i in ruchen_top2020_basic$id){
  ruchen_top2020_wfeat <- rbind(ruchen_top2020_wfeat, getFeatures(i, token = keys))
}
                  
# get rid of first line (that we added to set up the data frame)
ruchen_top2020_wfeat <- ruchen_top2020_wfeat[-1,]

# join playlist df and feature df to have all song info in one data set
ruchen_top2020 <- left_join(ruchen_top2020_basic, ruchen_top2020_wfeat, by = "id") 

# fix data set
ruchen_top2020 <- ruchen_top2020 %>% 
                  # eliminate irrelevant columns
                  select(-uri, -analysis_url, -time_signature) %>% 
                  # add 1 columns with name for later when joining all of us 
                  cbind("Ruchen") %>% 
                  # rename that column
                  rename(name = '"Ruchen"')

# check if there are duplicates
dupes <- ruchen_top2020 %>%
  get_dupes(id)

# get rid of duplicates
ruchen_top2020 <- ruchen_top2020 %>%
  unique()

# check
dupes <- ruchen_top2020 %>%
  get_dupes(id)

# check data 
#head(ruchen_top2020)
#glimpse(ruchen_top2020)
#skim(ruchen_top2020)
```

### Filippo

- original: https://open.spotify.com/playlist/37i9dQZF1EtsyZafU4Vjrk

```{r retrieve playlist Filippo}
# get playlist
filippo_top2020_basic <- getPlaylistSongs("spotify","37i9dQZF1EtsyZafU4Vjrk", token =  keys)

# get all features for that playlist

# create data set with any 1 song
filippo_top2020_wfeat <- getFeatures("3bH4HzoZZFq8UpZmI2AMgV", token = keys)

# for loop all song ids from the playlist through the getFeatures function
for (i in filippo_top2020_basic$id){
  filippo_top2020_wfeat <- rbind(filippo_top2020_wfeat, getFeatures(i, token = keys))
}
                  

# get rid of first line (that we added to set up the data frame)
filippo_top2020_wfeat <- filippo_top2020_wfeat[-1,]

# join playlist df and feature df to have all song info in one data set
filippo_top2020 <- left_join(filippo_top2020_basic, filippo_top2020_wfeat, by = "id") 

# fix data set
filippo_top2020 <- filippo_top2020 %>% 
                  # eliminate irrelevant columns
                  select(-uri, -analysis_url, -time_signature) %>% 
                  # add 1 columns with name for later when joining all of us 
                  cbind("Filippo") %>% 
                  # rename that column
                  rename(name = '"Filippo"')

# check if there are duplicates
dupes <- filippo_top2020 %>%
  get_dupes(id)

# get rid of duplicates
filippo_top2020 <- filippo_top2020 %>%
  unique()

# check
dupes <- filippo_top2020 %>%
  get_dupes(id)

# check data 
#head(filippo_top2020)
#glimpse(filippo_top2020)
#skim(filippo_top2020)

```

### Jason

- original: https://open.spotify.com/playlist/37i9dQZF1EM60SaMcpGdCv

```{r retrieve playlist Jason}
# get playlist
jason_top2020_basic <- getPlaylistSongs("spotify","37i9dQZF1EM60SaMcpGdCv", token =  keys)

jason_top2020_basic <- jason_top2020_basic %>%
  unique()
# get all features for that playlist

# create data set with any 1 song
jason_top2020_wfeat <- getFeatures("3bH4HzoZZFq8UpZmI2AMgV", token = keys)

# for loop all song ids from the playlist through the getFeatures function
for (i in jason_top2020_basic$id){
  jason_top2020_wfeat <- rbind(jason_top2020_wfeat, getFeatures(i, token = keys))
}
                  
# get rid of first line (that we added to set up the data frame)
jason_top2020_wfeat <- jason_top2020_wfeat[-1,]

# join playlist df and feature df to have all song info in one data set
jason_top2020 <- left_join(jason_top2020_basic, jason_top2020_wfeat, by = "id") 

# fix data set
jason_top2020 <- jason_top2020 %>% 
                  # eliminate irrelevant columns
                  select(-uri, -analysis_url, -time_signature) %>% 
                  # add 1 columns with name for later when joining all of us 
                  cbind("Jason") %>% 
                  # rename that column
                  rename(name = '"Jason"')

# check if there are duplicates
dupes <- jason_top2020_wfeat %>%
  get_dupes(id)

# get rid of duplicates
jason_top2020 <- jason_top2020 %>%
  unique()

# check
dupes <- jason_top2020 %>%
  get_dupes(id)

# check data 
#head(jason_top2020)
#glimpse(jason_top2020)
#skim(jason_top2020)

```


## Plot radar charts

```{r marie_rc}
marie_rc <- marie_top2020 %>% 
  select(danceability, valence, speechiness, instrumentalness, acousticness, liveness, energy) %>% 
  ggRadar(colour = "#8000FF",
          # transparency
          alpha = 0.8, 
          # get rid of points
          size = 0,
          # set y axis limits
          ylim = c(0,1))+
  labs(title = "Marie",
       x = "",
       y = "",
       caption = "Source: https://developer.spotify.com") +
  # edit theme
  theme(text = element_text(colour = "white", size = 10, family = "Helvetica Neue Bold"),
        plot.title = element_text(size = 15, family = "Helvetica Neue Bold", hjust = 0.5),
        plot.subtitle = element_text(size = 11, family = "Helvetica Neue Light"),
        plot.caption = element_text(size = 6, family = "Helvetica Neue Light"),
        axis.title.x = element_text(size = 10, family = "Helvetica Neue Bold"),
        axis.text.x = element_text(colour = "white", size = 10, family = "Helvetica Neue Medium"),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_line(colour = "#f2f2f2", 
                                  size = 0.15, 
                                  linetype = 2),
        plot.background = element_rect(fill = "#191414"),
        panel.background = element_rect(fill = "#191414"),
        legend.position = "none",
        )+
NULL
```

```{r ruchen_rc}
ruchen_rc <- ruchen_top2020 %>% 
  select(danceability, valence, speechiness, instrumentalness, acousticness, liveness, energy) %>% 
  ggRadar(colour = "#FF0080",
          # transparency
          alpha = 0.8, 
          # get rid of points
          size = 0,
          # set y axis limits
          ylim = c(0,1))+
  labs(title = "Ruchen",
       x = "",
       y = "",
       caption = "Source: https://developer.spotify.com") +
  # edit theme
  theme(text = element_text(colour = "white", size = 10, family = "Helvetica Neue Bold"),
        plot.title = element_text(size = 15, family = "Helvetica Neue Bold", hjust = 0.5),
        plot.subtitle = element_text(size = 11, family = "Helvetica Neue Light"),
        plot.caption = element_text(size = 6, family = "Helvetica Neue Light"),
        axis.title.x = element_text(size = 10, family = "Helvetica Neue Bold"),
        axis.text.x = element_text(colour = "white", size = 10, family = "Helvetica Neue Medium"),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_line(colour = "#f2f2f2", 
                                  size = 0.15, 
                                  linetype = 2),
        plot.background = element_rect(fill = "#191414"),
        panel.background = element_rect(fill = "#191414"),
        legend.position = "none")+
NULL
```

```{r sammy_rc}
sammy_rc <- sammy_top2020 %>% 
  select(danceability, valence, speechiness, instrumentalness, acousticness, liveness, energy) %>% 
  ggRadar(colour = "#ffff00",
          # transparency
          alpha = 0.8, 
          # get rid of points
          size = 0,
          # set y axis limits
          ylim = c(0,1))+
  labs(title = "Sammy",
       x = "",
       y = "",
       caption = "Source: https://developer.spotify.com") +
  # edit theme
  theme(text = element_text(colour = "white", size = 10, family = "Helvetica Neue Bold"),
        plot.title = element_text(size = 15, family = "Helvetica Neue Bold", hjust = 0.5),
        plot.subtitle = element_text(size = 11, family = "Helvetica Neue Light"),
        plot.caption = element_text(size = 6, family = "Helvetica Neue Light"),
        axis.title.x = element_text(size = 10, family = "Helvetica Neue Bold"),
        axis.text.x = element_text(colour = "white", size = 10, family = "Helvetica Neue Medium"),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_line(colour = "#f2f2f2", 
                                  size = 0.15, 
                                  linetype = 2),
        plot.background = element_rect(fill = "#191414"),
        panel.background = element_rect(fill = "#191414"),
        legend.position = "none")+
NULL
```

```{r filippo_rc}
filippo_rc <- filippo_top2020 %>% 
  select(danceability, valence, speechiness, instrumentalness, acousticness, liveness, energy) %>% 
  ggRadar(colour = "#00FF80",
          # transparency
          alpha = 0.8, 
          # get rid of points
          size = 0,
          # set y axis limits
          ylim = c(0,1))+
  labs(title = "Filippo",
       x = "",
       y = "",
       caption = "Source: https://developer.spotify.com") +
  # edit theme
  theme(text = element_text(colour = "white", size = 10, family = "Helvetica Neue Bold"),
        plot.title = element_text(size = 15, family = "Helvetica Neue Bold", hjust = 0.5),
        plot.subtitle = element_text(size = 11, family = "Helvetica Neue Light"),
        plot.caption = element_text(size = 6, family = "Helvetica Neue Light"),
        axis.title.x = element_text(size = 10, family = "Helvetica Neue Bold"),
        axis.text.x = element_text(colour = "white", size = 10, family = "Helvetica Neue Medium"),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_line(colour = "#f2f2f2", 
                                  size = 0.15, 
                                  linetype = 2),
        plot.background = element_rect(fill = "#191414"),
        panel.background = element_rect(fill = "#191414"),
        legend.position = "none")+
NULL
```

```{r jason_rc}
jason_rc <- jason_top2020 %>% 
  select(danceability, valence, speechiness, instrumentalness, acousticness, liveness, energy) %>% 
  ggRadar(colour = "#00FFFF",
          # transparency
          alpha = 0.8, 
          # get rid of points
          size = 0,
          # set y axis limits
          ylim = c(0,1))+
  labs(title = "Jason",
       x = "",
       y = "",
       caption = "Source: https://developer.spotify.com") +
  # edit theme
  theme(text = element_text(colour = "white", size = 10, family = "Helvetica Neue Bold"),
        plot.title = element_text(size = 15, family = "Helvetica Neue Bold", hjust = 0.5),
        plot.subtitle = element_text(size = 11, family = "Helvetica Neue Light"),
        plot.caption = element_text(size = 6, family = "Helvetica Neue Light"),
        axis.title.x = element_text(size = 10, family = "Helvetica Neue Bold"),
        axis.text.x = element_text(colour = "white", size = 10, family = "Helvetica Neue Medium"),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_line(colour = "#f2f2f2", 
                                  size = 0.15, 
                                  linetype = 2),
        plot.background = element_rect(fill = "#191414"),
        panel.background = element_rect(fill = "#191414"),
        legend.position = "none")+
NULL
```

```{r all rc}
#grid.arrange(filippo_rc, jason_rc, marie_rc, ruchen_rc, sammy_rc, nrow = 3)
filippo_rc
jason_rc
marie_rc
ruchen_rc
sammy_rc
```

```{r rc gif}
# save each chart as jpeg
filippo_rc + ggsave('filippo_rc.jpeg', height = 5, width = 5, units = 'in')
jason_rc + ggsave('jason_rc.jpeg', height = 5, width = 5, units = 'in')
marie_rc + ggsave('marie_rc.jpeg', height = 5, width = 5, units = 'in')
ruchen_rc + ggsave('ruchen_rc.jpeg', height = 5, width = 5, units = 'in')
sammy_rc + ggsave('sammy_rc.jpeg', height = 5, width = 5, units = 'in')
```


# Density charts

Next, we want to directly compare our music tastes.

## Data Preparation

First, we want to combine all our songs into one data set

```{r combined dataset}
# combine data sets
spotify_all <- marie_top2020 %>% 
                rbind(sammy_top2020) %>% 
                rbind(ruchen_top2020) %>% 
                rbind(filippo_top2020) %>% 
                rbind(jason_top2020)

# check data sets
glimpse(spotify_all)
skim(spotify_all)

```

Now, let's visualise our differences! (or similarities?)

## Tempo

```{r tempo}
levels <- c("Sammy","Ruchen", "Marie", "Jason","Filippo")
favcolours <- c("#ffff00", "#FF0080", "#8000FF", "#00FFFF", "#00FF80")
label <- ("Filippo likes songs\nwith the same tempo")



tempo <- spotify_all %>% 
  mutate(name = factor(name, levels = levels)) %>% 
  ggplot(aes(x = tempo, y = name, fill = name, colour = name))+
  geom_density_ridges(alpha = 0.4) +
  # add labels
  labs(title = "Not too fast and not too slow",
       subtitle = "Beats per minute (BPM) distribution of our favourite songs per person",
       x = "Tempo in BPM",
       y = "",
       caption = "Source: https://developer.spotify.com") +
  # edit colours
  scale_fill_manual(values = favcolours)+
  scale_colour_manual(values = favcolours)+
  # edit theme
  theme(text = element_text(colour = "white", size = 11, family = "Helvetica Neue Light"),
        plot.title = element_text(size = 15, family = "Helvetica Neue Bold"),
        plot.subtitle = element_text(size = 11, family = "Helvetica Neue Light"),
        plot.caption = element_text(size = 6, family = "Helvetica Neue Light"),
        axis.title.x = element_text(size = 10, family = "Helvetica Neue Medium"),
        axis.text.x = element_text(colour = "white"),
        axis.text.y = element_text(colour = "white", size = 12, family = "Helvetica Neue Bold"),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "#191414"),
        panel.background = element_rect(fill = "#191414"),
        plot.margin=grid::unit(c(0,0,0,0), "mm"),
        legend.position = "none")+
  # add descriptive text
   geom_text(data = data.frame(x = 50, y = 5.5, label = label),
          aes(x = x, y = y, label = label),
          colour="white",
          family="Helvetica Neue Medium",
          size = 3,
          inherit.aes = FALSE
          )+
   # add arrow
   geom_curve(aes(x = 60, y = 5.85, xend = 110, yend = 6.3),
              arrow = arrow(length = unit(0.04, "npc")),
              colour = "white",
              curvature = -0.4,
              size = 0.2)+
  NULL

tempo + ggsave('tempo.jpeg')
```


## Song popularity

```{r popularity}
levels <- c("Sammy","Ruchen", "Marie", "Jason","Filippo")
favcolours <- c("#ffff00", "#FF0080", "#8000FF", "#00FFFF", "#00FF80")
label2 <- ("Marie enjoys\nunknown songs")



popularity <- spotify_all %>% 
  mutate(name = factor(name, levels = levels)) %>% 
  ggplot(aes(x = popularity, y = name, fill = name, colour = name))+
  geom_density_ridges(alpha = 0.4) +
  # add labels
  labs(title = "How mainstream are our favourite songs?",
       subtitle = "Popularity distribution of our favourite songs per person",
       x = "Song popularity",
       y = "",
       caption = "Source: https://developer.spotify.com") +
  # edit colours
  scale_fill_manual(values = favcolours)+
  scale_colour_manual(values = favcolours)+
  # edit x axes
  scale_x_continuous(breaks = c(0,100),
                     labels = c("Unknown", "Most popular"))+
  # edit theme
  theme(text = element_text(colour = "white", size = 11, family = "Helvetica Neue Light"),
        plot.title = element_text(size = 15, family = "Helvetica Neue Bold"),
        plot.subtitle = element_text(size = 11, family = "Helvetica Neue Light"),
        plot.caption = element_text(size = 6, family = "Helvetica Neue Light"),
        axis.title.x = element_text(size = 10, family = "Helvetica Neue Bold"),
        axis.text.x = element_text(colour = "white", size = 10, family = "Helvetica Neue Medium"),
        axis.text.y = element_text(colour = "white", size = 12, family = "Helvetica Neue Bold"),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "#191414"),
        panel.background = element_rect(fill = "#191414"),
        plot.margin=grid::unit(c(1,1,1,1), "mm"),
        legend.position = "none")+
  # add descriptive text
   geom_text(data = data.frame(x = 0, y = 1.5, label = label2),
          aes(x = x, y = y, label = label2),
          colour="white",
          family="Helvetica Neue Medium",
          size = 3,
          inherit.aes = FALSE
          )+
   # add arrow
   geom_curve(aes(x = 11, y = 1.6, xend = 5, yend = 3.2),
              arrow = arrow(length = unit(0.04, "npc")),
              colour = "white",
              curvature = 0.4,
              size = 0.2)+
  NULL

popularity + ggsave('popularity.jpeg')
```

> But what makes a song popular? Surely the most popular playlist on Spotify will tell us!

# Song popularity

## Today's Top 100

We checked today's most popular playlist on Spotify and used the audio features mentioned above to find out why today’s top hits are the most popular.

## Distribution of song dimensions

Wanted to explore gender dynamics for artist > manually imputed gender dynamics onto playlist

This playlist is used: https://open.spotify.com/playlist/37i9dQZF1DXcBWIGoYBM5M

```{r top 100 dimension distribution}
popular <- read_csv("popular.csv")

popular<-janitor::clean_names(popular)

pp<-popular%>%
  select(id,valence,acousticness,danceability,energy,instrumentalness,liveness,speechiness,artist_gender)%>%
  pivot_longer(cols=!c(id,artist_gender), names_to="measure",values_to="count")%>%
  group_by(measure) %>%
  mutate(measure = fct_reorder(measure,count,.fun='median'))

box_pop <- ggplot(data=pp,aes(x=reorder(measure,desc(count)),y=count,color=as.factor(artist_gender), fill=as.factor(artist_gender)))+
  geom_half_boxplot(data=pp%>%filter(artist_gender=="Male"),side = "l")+
  geom_half_boxplot(data=pp%>%filter(artist_gender=="Female"),side = "r")+
scale_fill_manual(values = c("#191414","#191414"))+
scale_color_manual(values = c("#bc13fe","#1DB954"))+
  theme(text = element_text(colour = "white", size = 10, family = "Helvetica Neue Bold"),
        plot.title = element_text(size = 16, family = "Helvetica Neue Bold", hjust = 0),
        plot.subtitle = element_text(size = 11, family = "Helvetica Neue Light", hjust = 0),
        plot.caption = element_text(size = 6, family = "Helvetica Neue Light"),
        axis.title.x = element_text(size = 10, family = "Helvetica Neue Bold"),
        axis.text.x = element_text(colour = "white", size = 10, family = "Helvetica Neue Medium"),
        axis.text.y = element_text(colour = "white", size = 10, family = "Helvetica Neue Bold"),
        axis.line = element_line(linetype = 1, colour = "grey80"),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "#191414"),
        panel.background = element_rect(fill = "#191414"),
        legend.background =element_rect(fill = "#191414"),
        legend.box.background = element_rect(fill = "#191414", colour = "#191414"),
        legend.key = element_rect(fill = "#191414"),
        legend.position = "top"
        )+
  scale_y_continuous(breaks = c(0,1),
                     labels = c("0%", "100%"))+
  labs(title = "Popular Music is all about Danceability and Energy", 
       subtitle = "No one cares about the instrument you play",
       x = "",
       y = "",
       colour="",fill="")

box_pop + ggsave('boxplot_of_top_songs.jpeg', units = 'in')

```

> But this is only what makes today’s music popular. Let’s see which factors are most important in determining popularity!

## Regression

We want to find out which features play a key part in determining popularity on a more representative dataset.

The "data_all.csv" file contains more than 170.000 songs collected from Spotify Web API. Each row of data contains the following information:

- id (Id of track generated by Spotify)
- acousticness (Ranges from 0 to 1)
- danceability (Ranges from 0 to 1)
- energy (Ranges from 0 to 1)
- duration_ms (Integer typically ranging from 200k to 300k)
- instrumentalness (Ranges from 0 to 1)
- valence (Ranges from 0 to 1)
- popularity (Ranges from 0 to 100)
- tempo (Float typically ranging from 50 to 150)
- liveness (Ranges from 0 to 1)
- loudness (Float typically ranging from -60 to 0)
- speechiness (Ranges from 0 to 1)
- year (Ranges from 1921 to 2020)
- mode (0 = Minor, 1 = Major)
- explicit (0 = No explicit content, 1 = Explicit content)
- key (All keys on octave encoded as values ranging from 0 to 11, starting on C as 0, C# as 1 and so on…)
- artists (List of artists mentioned)
- release_date (Date of release mostly in yyyy-mm-dd format, however precision of date may vary)
- name (Name of the song)

Based on this dataset we run a regression analysis with popularity as the dependent variable.

```{r regression analysis}
data<- read_csv("data_all.csv")
View(data)

set.seed(1)
library(rsample)
train_test_split <- initial_split(data, prop = 0.8) #training set contains 75% of the data
# Create the training dataset
train_data <- training(train_test_split)
test_data <- testing(train_test_split)

control <- trainControl (
  method="cv",
  number=5,
  verboseIter=TRUE) #by setting this to true the model will report its progress after each estimation


#we are going to train the model and report the results using k-fold cross validation
model1_lm<-train(
  (popularity)~valence+acousticness+danceability+duration_ms+energy+explicit+instrumentalness+liveness+loudness+speechiness+tempo,
  train_data,
  method = "lm",
  trControl = control
)


summary(model1_lm)

importance <- varImp(model1_lm, scale=TRUE)

# create feature importance map

# saved importance feature values
varImp <- read_csv("varImp.csv")

skimr::skim(varImp)

varImp%>%
  arrange(importance) %>%
  mutate(type = factor(type, levels=c("duration_ms","tempo", "loudness","liveness","energy","explicit","instrumentalness","danceability","valence","speechiness","acousticness"))) %>%ggplot(aes(x=reorder(type,desc(importance)),y=importance))+
  geom_segment(aes(x=type,xend=type,y=0,yend=importance),color="green")+geom_point(size=4, color="green")+
  coord_flip() +
  theme(text = element_text(colour = "white", size = 10, family = "Helvetica Neue Bold"),
        plot.title = element_text(size = 16, family = "Helvetica Neue Bold", hjust = 0),
        plot.subtitle = element_text(size = 11, family = "Helvetica Neue Light", hjust = 0),
        plot.caption = element_text(size = 6, family = "Helvetica Neue Light"),
        axis.title.x = element_text(size = 10, family = "Helvetica Neue Bold"),
        axis.text.x = element_text(colour = "white", size = 10, family = "Helvetica Neue Medium"),
        axis.text.y = element_text(colour = "white", size = 10, family = "Helvetica Neue Bold"),
        axis.line = element_line(linetype = 1, colour = "grey80"),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "#191414"),
        panel.background = element_rect(fill = "#191414"),
        legend.background =element_rect(fill = "#191414"),
        legend.box.background = element_rect(fill = "#191414"),
        legend.key = element_rect(fill = "#191414"),
        legend.position = "top")+
  labs(title = "What makes a song popular?", 
       subtitle = "No one cares about the instrument you play",
       x = "",
       y = "Importance",
       colour="",fill="")
```


> Unfortunately, song popularity isn’t this simple. We have to take more factors into consideration… 

> Content, genres, time & location

## Content

In this section we use text mining on the top 50 World Playlists to understand what are the most frequently appeared words.
             

```{r by content analysis}
# Construct path to a folder where our songs' corpus will be built
yourPath <- "Lyrics_groupproject"
name <- file.path(yourPath)
length(dir(name)) 
dir(name) 

# Creating corpus 
# tm package function: store corpus into variable called 'original' 
original <- Corpus(DirSource(name)) 
# create a new variable called docs
docs <- original 

# Short summary of Corpus
typeof(docs) # "list"
class(docs)  # "SimpleCorpus" "Corpus"      

length(docs) # 48
show(docs)
# <<SimpleCorpus>>
# Metadata:  corpus specific: 1, document level (indexed): 0
# Content:  documents: 48
summary(docs)
#             Length Class             Mode
# 34+35 - Ariana grande.txt                      2      PlainTextDocument list
# ALWAYS DO - The Kid Laroi.txt                  2      PlainTextDocument list
# Angels Like You - Miley Cyrus.txt              2      PlainTextDocument list
# At My Worst - Pink Sweat$.txt                  2      PlainTextDocument list
# Bichota - KAROL F.txt                          2      PlainTextDocument list
# Blinding Lights - The Weeknd.txt               2      PlainTextDocument list
# Body - Megan Thee Stallion.txt                 2      PlainTextDocument list
# Breaking - Topic.txt                           2      PlainTextDocument list
# many more...


# Inspecting documents
docs[[1]]$content # 1st doc' actual word content
docs[[1]]$meta    # additional meta data which gets created during analysis (we won't need this)


# Transformations of corpus text 
docs <- tm_map(docs, tolower) # tm package function
inspect(docs[[1]])

docs <- tm_map(docs, removeNumbers)
inspect(docs[[1]])

stopwords("english")
docs <- tm_map(docs, removeWords, stopwords("english"))
inspect(docs[[1]])

docs <- tm_map(docs, removePunctuation)
inspect(docs[[1]])

docs <- tm_map(docs, stripWhitespace)
inspect(docs[[1]])

words <- c("con", "dadadadadada","lalalala","tã°", "pumpum", "que", "get", "let", "one", "two", "three", "four","can","got","cause","now","just")
docs <- tm_map(docs, removeWords, words)
inspect(docs[[1]])


# Examine 1st document CONTENT after transformations
text1 <- docs[[1]]$content 
# Let's check the data type of this variable
typeof(text1) # 'character' 
length(text1) # 1 i.e. it is a single string of all words in that document
print(text1)
class(docs)      # "SimpleCorpus" "Corpus"      
class(docs[[1]]) # "PlainTextDocument" "TextDocument"   

# Create Document Term Matrix 
# Use term frequency (TF) to create the matrix
library(tm)
DTM <- DocumentTermMatrix(docs) 
dim(DTM) # 48 (docs / rows) x 2167 (words / cols) 
print(DTM)
# <<DocumentTermMatrix (documents: 48, terms: 2167)>>
# Non-/sparse entries: 3750/100266
# Sparsity           : 96% 
# Maximal term length: 32
# Weighting          : term frequency (tf)
inspect(DTM[, 1:5]) 

# Convert to tibble
m <- as.matrix(DTM)
# from=x, to=y, by=z i.e. in steps of 1
year <- seq(0, 100, 1) 
# bind column called year, and other columns in matrix into a single matrix M
M <- cbind(year, m) 
# transform matrix into tibble
DTM_tbl <- as_tibble(m) 
DTM_tbl_tidy <- pivot_longer(DTM_tbl, cols = !year, names_to = "word", values_to = "wordCount")

# Analysis
# 1) total count
# number of words in each doc
# Tidyverse approach
totCount <- DTM_tbl_tidy %>%  
        group_by(year) %>% 
        summarise(total = sum(wordCount)) %>% 
        arrange(desc(total)) 
print(totCount)


# Base R approach:
tot <- rowSums(as.matrix(DTM))
totCount <- sort(tot, decreasing = TRUE) 
print(totCount)

# 2) Word frequency 
# Number of times each word occurred across all documents
# Tidyverse approach
wordCountDoc <- DTM_tbl_tidy %>%  
        group_by(word) %>% 
        summarise(total = sum(wordCount)) %>% 
        arrange(desc(total)) 
print(wordCountDoc)


# Base R approach:
freq <- colSums(as.matrix(DTM)) 
wordCountDoc2 <- sort(freq, decreasing = TRUE) 
print(wordCountDoc2[1:10]) # obtain a slice of vector 'wordCountDoc' starting from element 1 to 10

# 3) Word frequency table 
wordFreqTable <- table(wordCountDoc2)
head(wordFreqTable) 
tail(wordFreqTable)


# 4) N word occurances 
# Find which words occurred at least 125 times
top_n <- findFreqTerms(DTM, 125)
print(top_n)
# "know" "like" "yeah"

# 5) Bar Chart of n most freq words

DTM_tbl_tidy %>%  
        group_by(word) %>% 
        summarise(total = sum(wordCount)) %>% 
        arrange(desc(total)) %>% 
        top_n(10) %>% 
        ggplot(aes(x = reorder(word, total), y = total, fill = word)) +
        geom_bar(stat = "identity") + 
        theme(legend.position = "none",
              axis.text = element_text(size = 10, colour= "green"),
              plot.title = element_text(size = 10, face = "bold", colour = "green",hjust = 0.5),
              axis.title = element_text(size = 10), plot.subtitle = element_text(size = 8,hjust = 0.5, colour = "green"),panel.background = element_rect(fill = "black",
                                                                                                                                                         colour = "black",
                                                                                                                                                         size = 0.5, linetype = "solid"),plot.background = element_rect(fill = "black")) +
        coord_flip() +
        labs(title = "Top 10 Most Frequent Words in the
Spotify Top 50 World Playlist",
             x = "",
             y = "", subtitle = "excluding common verbs and conjunctions")



# word clouds

# figPaths = system.file("spotify.png",package = "wordcloud2")
wordcloud2(wordCountDoc, color = "random-light",backgroundColor = "black") 

```


## Genres

Music preferences differ highly by genre, which somehow corresponds to listens' mood or social identity. Here we choose 6 most identifiable genres to see the feature distribution using radar charts. 


########## WE WILL ADD CODE BUT IT'S BASICALLY MORE RADAR CHARTS BUT FOR 6 GENRES

## Time

We use the kaggle dataset and build an interactive story consisting of three dashboards.
The first one visualizes the top 10 prolific artists in each decade. 1950s to 1980s have witnessed proliferation of artist creations (Queen, The Beatles, The Rolling Stones, Bob Dylan).

The second line charts shows how the audio features about mood, properties, and context are changing over time. On average, music is getting louder, more energetic, and more explicit. Meanwhile, valence, instrumentalness and speechiness are also decreasing. 

The relevant artists and tracks in each decade rated from today's audience are changing. The Beatles and Led Zeppelin are audience’ all time favorites. Pop Smoke has two tracks in the top 10 popular tracks. 

Feel free to play with different filters more [here](https://public.tableau.com/profile/ruchen8489#!/vizhome/Group14_Spotify_ByDecade/Story?publish=yes)



## Location

We use the data set containing top hit playlists of countries around the globe and built a dashboard consisting of four plots.

The "data_by_country.csv" file contains top 200 tracks figures(track name, artist and streaming) per day per country around the globe. You can download csv [here](https://spotifycharts.com/regional)

You can click on the country you are interested in on the map and the other three plots change immediately to show daily steaming trend change, most popular tracks and artists in that specific country you've chosen!

We found that steaming trends rise globally. Countries on the same continent share similar tastes except Ed Sheeran and Ariana Grande conquer the world. Meanwhile, countries speaking different languages tend to have different favorite songs

It is interesting to see that one popular song is sometimes not enough to make the artist popular. For example, in Japan, Kenshi Yonezu is not even ranked among top 10 artists, but one of his songs is ranked 2# most popular. While popular artists always publish most popular tracks. In U.K., top 2 songs are produced by top 1 artist.

Find out more [here](https://public.tableau.com/profile/sammy.chen#!/vizhome/Musictrendandtasteanalysisbycountry/Dashboard1?publish=yes)

# Conclusion

**There is no one way to understand and predict music preferences and popularity**

* However similar they are, **every individual’s music taste is unique.**

* **Today’s popular music all about danceability and energy;** it is less about live music, acoustic tracks or use of actual instruments

* Acoustic songs, conveying positive moods or using a lot of speech seem to negatively influence song popularity while danceable songs with high energy and explicit words seem to guarantee success - Caution though! **Music is too diverse to be generalised in a statistical model**

* We must keep the following in mind

  1. *Music preferences differ by genre*
  2. *Lyrics and their meaning highly influence songs*
  3. *Music taste and preferences evolve over time;* however, some artists and songs survive time and are still relevant today
  4. Either way, *success is often short* and all artists try to survive in a highly competitive music industry
  5. *Music taste and preferences differ by country;* some songs reach people worldwide while others stay local

Everyone’s music taste is unique.

You can explore the data as much as you like – at the end of the day, only you know what makes you like a song!


















## Indie

### Playlist

- original: https://open.spotify.com/playlist/37i9dQZF1DX2Nc3B70tvx0

Retrieve playlist through API, then retrieve the features of every song with a for loop and create a data set with all the songs and their features

```{r retrieve playlist indie}
# get playlist
indie_basic <- getPlaylistSongs("spotify","37i9dQZF1DX2Nc3B70tvx0", token =  keys)

indie_basic <- indie_basic %>%
  unique()
# get all features for that playlist

# create data set with any 1 song
indie_wfeat <- getFeatures("3bH4HzoZZFq8UpZmI2AMgV", token = keys)

# for loop all song ids from the playlist through the getFeatures function
for (i in indie_basic$id){
  indie_wfeat <- rbind(indie_wfeat, getFeatures(i, token = keys))
}
                  
```

Fix playlist data set and delete duplicates

```{r fix data indie}
# get rid of first line (that we added to set up the data frame)
indie_wfeat <- indie_wfeat[-1,]

# join playlist df and feature df to have all song info in one data set
indie <- left_join(indie_basic, indie_wfeat, by = "id") 

# fix data set
indie <- indie %>% 
                  # eliminate irrelevant columns
                  select(-uri, -analysis_url, -time_signature) %>% 
                  # add 1 columns with name for later when joining all of us 
                  cbind("Indie") %>% 
                  # rename that column
                  rename(name = '"Indie"')
```

```{r clean data indie}
# check if there are duplicates
dupes <- indie_wfeat %>%
  get_dupes(id)
```

### Radar chart

```{r indie rc}
indie_rc <- indie %>% 
  select(danceability, valence, speechiness, instrumentalness, acousticness, liveness, energy) %>% 
  ggRadar(colour = "#00FFFF",
          # transparency
          alpha = 0.8, 
          # get rid of points
          size = 0,
          # set y axis limits
          ylim = c(0,1))+
  labs(title = "Ultimate Indie",
       subtitle = "101 songs, 6h 21 min, 2,036,610 likes*",
       x = "",
       y = "",
       caption = "*Note: retrieved 02/12/2020\nSource: https://developer.spotify.com") +
  # edit theme
  theme(text = element_text(colour = "white", size = 10, family = "Helvetica Neue Bold"),
        plot.title = element_text(size = 16, family = "Helvetica Neue Bold", hjust = 0.5),
        plot.subtitle = element_text(size = 11, family = "Helvetica Neue Light", hjust = 0.5),
        plot.caption = element_text(size = 6, family = "Helvetica Neue Light"),
        axis.title.x = element_text(size = 10, family = "Helvetica Neue Bold"),
        axis.text.x = element_text(colour = "white", size = 10, family = "Helvetica Neue Medium"),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_line(colour = "#f2f2f2", 
                                  size = 0.15, 
                                  linetype = 2),
        plot.background = element_rect(fill = "#191414"),
        panel.background = element_rect(fill = "#191414"),
        legend.position = "none")+
NULL
```

## Hip Hop

- original: https://open.spotify.com/playlist/37i9dQZF1DWT5MrZnPU1zD

Retrieve playlist through API, then retrieve the features of every song with a for loop and create a data set with all the songs and their features

```{r hip hop}
# get playlist
hiphop_basic <- getPlaylistSongs("spotify","37i9dQZF1DWT5MrZnPU1zD", token =  keys)

hiphop_basic <- hiphop_basic %>%
  unique()
# get all features for that playlist

# create data set with any 1 song
hiphop_wfeat <- getFeatures("3bH4HzoZZFq8UpZmI2AMgV", token = keys)

# for loop all song ids from the playlist through the getFeatures function
for (i in hiphop_basic$id){
  hiphop_wfeat <- rbind(hiphop_wfeat, getFeatures(i, token = keys))
}
                  
# get rid of first line (that we added to set up the data frame)
hiphop_wfeat <- hiphop_wfeat[-1,]

# join playlist df and feature df to have all song info in one data set
hiphop <- left_join(hiphop_basic, hiphop_wfeat, by = "id") 

# fix data set
hiphop <- hiphop %>% 
                  # eliminate irrelevant columns
                  select(-uri, -analysis_url, -time_signature) %>% 
                  # add 1 columns with name for later when joining all of us 
                  cbind("Hip Hop") %>% 
                  # rename that column
                  rename(name = '"Hip Hop"')

# check if there are duplicates
dupes <- hiphop_wfeat %>%
  get_dupes(id)

hiphop_rc <- hiphop %>% 
  select(danceability, valence, speechiness, instrumentalness, acousticness, liveness, energy) %>% 
  ggRadar(colour = "#00FFFF",
          # transparency
          alpha = 0.8, 
          # get rid of points
          size = 0,
          # set y axis limits
          ylim = c(0,1))+
  labs(title = "Hip Hop Controller",
       subtitle = "100 songs, 5h 4 min, 1,373,660 likes*",
       x = "",
       y = "",
       caption = "*Note: retrieved 02/12/2020\nSource: https://developer.spotify.com") +
  # edit theme
  theme(text = element_text(colour = "white", size = 10, family = "Helvetica Neue Bold"),
        plot.title = element_text(size = 16, family = "Helvetica Neue Bold", hjust = 0.5),
        plot.subtitle = element_text(size = 11, family = "Helvetica Neue Light", hjust = 0.5),
        plot.caption = element_text(size = 6, family = "Helvetica Neue Light"),
        axis.title.x = element_text(size = 10, family = "Helvetica Neue Bold"),
        axis.text.x = element_text(colour = "white", size = 10, family = "Helvetica Neue Medium"),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_line(colour = "#f2f2f2", 
                                  size = 0.15, 
                                  linetype = 2),
        plot.background = element_rect(fill = "#191414"),
        panel.background = element_rect(fill = "#191414"),
        legend.position = "none")+
NULL
```



```{r all genres}
indie_rc
hiphop_rc

indie_rc + ggsave('indie.jpeg', height = 5, width = 5, units = 'in')
```












